\section{Conclusion} \label{sec:conclusion}

DBMS performance tuning is a niche skill that requires much experience
and experimentation to get right. Machine learning techniques can help 
automate this process by using prior knowledge to estimate the 
performance of the DBMS on a particular SQL workload 
without having to actually execute it. 
To achieve this goal, we have taken a two-step approach where 
we first map a workload to a well-known benchmark. 
Armed with this information, we then use a benchmark-specific
regression estimator to obtain performance estimates 
under different DBMS configurations.

Using decision trees, we find that we can classify workloads 
using a few defining features.  
Our experiments show that a CART-based decision tree classifier
achieves 91\% accuracy on the classification problem.
Thus, the decision trees learned to effectively discriminate 
between the different benchmark classes.
We use Gaussian Process Regression estimators for predicting 
the throughput and latency of the DBMS while executing a
SQL workload under a particular configuration.
We demonstrate that this estimator achieves a R-squared score 
of 0.9. 
Furthermore, these highly accurate estimators can be trained 
with as few as 600 samples. 
Our analysis using Lasso Regression estimators provided key 
insights into the features that are most influential in 
determining throughput and latency.

In our experiments, we generate synthetic variants of the
benchmarks in OLTPBench to emulate real-world workloads as
we felt that it is a good starting point for evaluating the
viability of our techniques.
We plan to continue our evaluation using a larger real-world
dataset that includes more realistic samples. 
We note that we leave the problem of using the performance 
estimator to find an optimal DBMS configuration for a given 
workload for future work. The estimator allows us to obtain 
inexpensive estimates without actually running the workload 
under a specific configuration.
In addition, we will extend our auto-tuning framework to 
support other DBMSs like MySQL, and more hardware profiles. 
We anticipate that this larger and more realistic dataset 
will help us obtain additional insights about how best to 
tune the DBMS configuration to maximize performance on 
arbitrary SQL workloads.
